### 2021-08-31 paper.Rmd
through page 15

---
author:
  - name: Stuart Lee
    affiliation: Monash University
    # use this syntax to add text on several lines
    address: |
      | Department of Econometrics and Business Statistics, Monash University
    email: stuart.a.lee@monash.edu
    orcid: 0000-0003-1179-8436
  - name: Dianne Cook*
    address: |
      | Department of Econometrics and Business Statistics, Monash University
    affiliation: Monash University
    email: dicook@monash.edu
    orcid: 000-0002-3813-7155
  - name: Natalia da Silva
    address: |
      | Instituto de Estadística (IESTA), Universidad de la República
    affiliation: Universidad de la República
    email: natalia@iesta.edu.uy
    orcid: 0000-0002-6031-7451
  - name: Ursula Laa 
    affiliation: University of Natural Resources and Life Sciences
    address:  |
      | Institute of Statistics, University of Natural Resources and Life Sciences
    email:  ursula.laa@boku.ac.at
    orcid: 0000-0002-0249-6439
  - name: Nicholas Spyrison
    address: |
      | Faculty of Information Technology, Monash University
    affiliation: Monash University
    email: nicholas.spyrison@monash.edu
    orcid: 0000-0002-8417-0212
  - name: Earo Wang
    address: |
      | Department of Statistics, The University of Auckland
    affiliation: The University of Auckland
    email: earo.wang@auckland.ac.nz
    orcid: 0000-0001-6448-5260
  - name: H. Sherry Zhang
    address: |
      | Department of Econometrics and Business Statistics, Monash University
    affiliation: Monash University
    email: huize.zhang@monash.edu
    orcid: 0000-0002-7122-1463
title:
  formatted: "Advanced Review: The State-of-the-Art on Tours for Dynamic Visualization of High-dimensional Data"
abstract: > 
  This article discusses a high-dimensional visualization technique called the tour, which can be used to view data in more than three dimensions. We review the theory and history behind the technique, as well as modern software developments and applications of the tour that are being found across the sciences and machine learning. 
keywords:
  formatted: ["tours", "data visualization", "high-dimensional data", "data science", "exploratory data analysis"]
output: 
  bookdown::pdf_document2:
    template: template/wires_template_rmd.tex
    keep_tex: true
    toc: false
    citation_package: biblatex
bibliography: "references.bib"
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  fig.show = 'hold',
  dpi = 300,
  cache = FALSE
  # out.width = "\\textwidth",
  # fig.path = here::here('img/')
)
```

<!--
Writing style: third person (it was done, ...)
-->

# Introduction

<!-- Purpose/why and what is high-d data (Di)-->

Data commonly arrives with more than two measured variables, which makes it more complicated to plot on a page. With multiple variables, especially if there is some association between variables, this would be called multivariate or high-dimensional data. When the variables are all numeric, or quantitative, visualization often relies on some form of dimension reduction. This can be done by taking linear projections, for example, principal component analysis (PCA) [@Hotelling1933-of], linear discriminant analysis (LDA) [@Fisher1936] or projection pursuit (PP) [@Friedman1974-ck]. It is also common to reduce dimension with nonlinear techniques like multidimensional scaling (MDS) [@Kruskal1964-do],  t-Distributed Stochastic Neighbor Embedding (t-SNE) [@Van_der_Maaten2008-qa] or Uniform Manifold Alignment and Projection (UMAP) [@mcinnes2020umap].

The term "high-dimensional" here refers to the  dimensionality of the Euclidean space. Figure \ref{fig:cubes} shows a way to imagine this. It shows a sequence of cube wireframes, ranging from one-dimensional (1D) through to five-dimensional (5D), where beyond 2D is a linear projection of the cube. As the dimension increases, a new orthogonal axis is added. For cubes, this is achieved by doubling the cube: a 2D cube consists of two 1D cubes, a 3D cube consists of two 2D cubes, and so forth.

\begin{figure*}[!h]
\centerline{\includegraphics[width=0.95\textwidth]{figures/cubes.png}}
\caption{Illustrating what is meant by "high-dimensional" in this paper, and a "linear projection". From a sequence of increasing dimension cubes, from 1D to 5D, as wireframes, it can be seen that as the dimension increase by one, the cube doubles. }
\label{fig:cubes}
\end{figure*}

The focus of this review will be on visualizing high-dimensional numerical data using linear projections, in particular, as provided by the grand tour [@Asimov1985-xr; @Buja2005-cx] and its recent derivatives. The reason being that it is not feasible to adequately review the very large area of visualizing high-dimensions, and there have been numerous developments in tours recently. An overview of the technique, and new modifications is provided, along with how these techniques can be used in a variety of applications.

A tour can be considered to be a dynamic graphic, because it shows a smooth sequence of projections over time, ideally with controls that allow stopping, reversing, changing direction, or going forward again. It can be useful to embed a tour into an interactive graphics system, where plots can be queried and elements highlighted e.g. GGobi [@Swayne2003] or XLispStat [@Tierney1991]. To create the smooth sequence, a geodesic interpolation is computed between consecutive frames. A geodesic interpolation produces movement along the shortest path between two planes. It allows the viewer to extrapolate from the low-dimensional shapes to a corresponding multivariate distribution, and is particularly useful for detecting clusters, outliers and non-linear dependence.

While tours are invaluable for assessing the geometry of data, they are by no means the only technique available for visualizing structure in high dimensional data. Very similar to a tour was the Andrews plot [@andrews] which uses a fourier series to create a smooth series of 1-D projections. But as @wegman-solka point out this is a pseudo-grand tour because it is not space-filling. @Moustafa2009 explains the relationship between a grand tour and the classical multivariate plot, the Andrews curve. An early technique proposed for assessing pairwise relationships between variables is the scatterplot matrix (SPLOM) [@Tukey1981-ov; @Chambers1983-dh; @Tukey1983-fj; @Carr1984-xw; @Becker1987-dw]. The SPLOM allows the viewer to assess correlation structure but does not scale to large numbers of variables. Similarly, parallel coordinates plots (PCP) can be used to explore correlation and collinearity [@Inselberg1985-kf; @Wegman1990-dy]. By placing multiple variables side-by-side in a PCP higher order, structure like clustering or lower dimensional embeddings are revealed. However the ordering of variables along the axis changes what can be learned, and @hurley-oldford suggest repeating variables to alleviate this. Another display that relies on variable ordering is the heatmap, which is widely used to visualize cluster structure in bioinformatics. @Wilkinson2009-pj provides a comprehensive history of this display in the social and natural sciences.

All of the aforementioned techniques can be enhanced through the use of interactivity. By combining views with interactive elements like tool tips or highlighting, the analyst is able to quickly interrogate interesting features of the data. One particularly important interaction technique in the history of statistical graphics is called brushing [@Becker1987-dw]. When brushing an analyst drags their mouse over the view, which results in a region (the brush) being drawn onto the canvas. When there are multiple views present the act of brushing can be thought of as a database query; points that fall inside the brush can be used to highlight or filter data on adjacent views (Figure \ref{fig:brushing}). This brushing technique is particularly useful when combined with the tour (Section \@ref(sec:spin)).

\begin{figure*}[!ht]
   \centering
         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-query.pdf}
         \includegraphics[width=0.45\textwidth]{figures/ggobi_book/brush-response.pdf}
\caption{When a user brushes data on the left panel, points that fall inside the region are highlighted in orange. The view on the right responds by highlighting the corresponding points. Adapted from \cite{Cook2007-be} Figure 2.12. The left plot shows a jittered scatterplot of two species and two sexes of crabs from \cite{CM74}, and the right plot shows two of their five physical measurements.}
\label{fig:brushing}
\end{figure*}

<!--Tours can be considered to be a dimension reduction technique. Visualizing high-dimensions often starts with dimension reduction, for example, principal component analysis (PCA) [@pearson-pca;@hotelling-pca;@joliffe-cadima-pca], which is a linear dimension reduction technique. PCA finds the linear combination (without the normality constraint of a projection vector) of maximum variance, and can be considered to be a special case of projection pursuit (PP) [@Friedman1974-ck] which maximizes an arbitrary function.-->

Nonlinear dimension reduction with the techniques t-SNE and UMAP, have become very popular in recent years, primarily for the ability to capture cluster structure in a succinct visual summary. However, it is only a summary, and it likely involves substantial warping of the original data space. Using the tour along with these techniques can illuminate the nature of the warped space, and reveal other structure lost in the dimension reduction.  Figure \ref{fig:tsne-tour} illustrates the difference in what can be learned from a tour in comparison with nonlinear dimension reduction using t-SNE. The 10-dimensional data comes from @Rauber2009-kk. The t-SNE view (plot A) shows six clearly separated clusters, all spherical, which have different sizes. The tour shows that the clusters do not look like this in the full data space. The clusters are at various distances apart and very different sizes, as can be seen from the four projections from a tour. The two green clusters are large and almost spherical, and far from the orange clusters. The orange clusters have one larger one, and three smaller, very close to each other. All of these are elliptical, which means that they have very little variability, actually no variability if you watch the full tour, in some of the 10 dimensions. This gives some deeper perspective to what is learned from t-SNE, and illustrates what the t-SNE dimension reduction has done: it has found small gaps between points and expanded these gaps to yield the representation. It should be noted, though, that the methods (t-SNE and tour) complement each other. The t-SNE view gives a clear indication of six clusters, which may have been overlooked on initial viewing with a tour. With this information is the invitation to look closer at the data in the tour, to see, that yes, indeed, there are three, tiny, tiny clusters very close to each other.

\begin{figure*}[!t]
   \centering
         \includegraphics[width=0.29\textwidth]{figures/cluster_example.pdf}
         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom2/1.jpg}
         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom2/5.jpg}
         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom2/3.jpg}
         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom3/1.jpg}
         \includegraphics[width=0.29\textwidth]{figures/cluster_example_zoom3/2.jpg}
     %\vfill
   %\begin{minipage}{\linewidth}     
   %\centering
   %      \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour1.png}
   %      \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour2.png}
  %       \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour3.png}
   %      \includegraphics[width=0.24\textwidth]{figures/cluster_example_new_tour4.png}
  % \end{minipage}%
\caption{Comparison of structure perception between nonlinear dimension reduction using t-SNE (A) and linear dimension reduction using a tour (B) from 10D simulated data. There are six clusters visible in the t-SNE view, three large and three small, roughly equidistant. The relative distance between the clusters in the original space is extremely varied, as can be seen from the selection of tour projections shown. The two green clusters are (almost) spherical in shape, because they are spherical in every projection viewed. They are also far from all four orange clusters. Three of the orange clusters are tiny, and very close to each other (just visible in B.1, and clearer when zoomed in B.5). All orange clusters are elliptical. The tour provides a more accurate rendering of the clusters in the high-dimensional space, and complements what is learned from the nonlinear dimension reduction.}
\label{fig:tsne-tour}
\end{figure*}

The rest of the review is structured as follows: Section \@ref(notation) defines the notation and components of a tour displays. By its nature a tour is most effective to analyst when combined with interactivity; Section \@ref(interaction) reviews the components of user interfaces for manipulating tour views, including manual tours (Section \@ref(tours)) which are useful to test structure of selected features. The implementation of tour paths in statistical software is reviewed in Section \@ref(software). Section \@ref(applications) shows the diverse applications of the tour in the natural sciences, machine learning, and applied statistics. Finally, Section \@ref(discussion) discusses future research directions for tours.

# Tours for high dimensional visualization {#notation}

## Notation

When using a tour, a sequence of $d$-dimensional linear projections are obtained from a $p$-dimensional space, where $d \ll p$. Let $X_{n \times p}$ be the data matrix, consisting of $n$ observations and $p$ variables, whose projections are of interest. A projection basis, $A_{p \times d}$, is a matrix that characterizes the direction from which the data are projected.  It needs to satisfy an orthonormality condition, which requires that all the columns in $A$ has unit length and are perpendicular to each other. With a data matrix and a projection basis, a projection of the data can be defined as $Y = X \cdot A$. The Huber plots and selected 1D projections shown in Figure \ref{fig:notation-huber} are a nice way to imagine the tour infrastructure. (The figures produced with a modified function from the R [@R_Core_Team2020-lr] package PPtreeViz [@Lee2018-gy], and are motivated by @Lee2005-gl Figure 1.) Two (from four) variables and the species labels from the palmer penguins data [@penguins] are used for illustration. In the Huber plots (left) [@huber1990] a scatterplot of the two variables coloured by species is displayed in the center, and the solid line represents the value of a PP index, computed for all 1D projections. (The dashed circle is a guide to help perceive high and low PP index values.) The line segments indicate the 1D data projection which is plotted in the density plot to the right. The density plot in the top row exhibits clustering of the three species, and corresponds to a 1D projection with a high PP index value. The density plot in the bottom row corresponds to a 1D projection with low PP index value, and can be seen to have a less interesting pattern.

<!-- SL: I think there needs to be more detail here to link to the next paragraph and more explanation about what the huber plot is showing -->

\begin{figure*}
   \centering
         \includegraphics[width=\textwidth]{figures/notation-huber1.pdf}
         \includegraphics[width=\textwidth]{figures/notation-huber2.pdf}
\caption{Huber's plot (left plots) of the penguins data (bill length and depth only). The solid line represents how well the projection onto different directions, separates the three species, and the dashed circle is a reference guide. The density plots (right plots) is the 1D projection of the data onto the direction outlined as the solid line in the Huber's plot. These plots are made using the R package PPtreeViz.}
\label{fig:notation-huber}
\end{figure*}

<!-- Figure 4:   Huber's plot (left plots) of the penguins data....Maybe will be better to  capital letter for axis labels (Count, Group and Best projected data). And maybe use the same Dark2 palette. -->

\begin{figure*}
   \centering
         \includegraphics[width=\textwidth]{figures/tour_path.png}
\caption{Illustration of a tour path, indicating target planes, and interpolation planes. Adapted from \cite{Buja2005-cx}, Figure 1.}
\label{fig:tour-path}
\end{figure*}

## Different target plane methods produce different tour types

<!-- SL: I think this needs to be linked up to the notation section -->

A tour path, that is a sequence of projections, can be generated by geodesic interpolation between a set of target planes (Figure \ref{fig:tour-path}). Different methods for choosing target planes provide different tour paths. The grand tour [@Asimov1985-xr] is generated using randomly selected targets. A guided tour [@Cook1995-ae] generates targets as a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$. A local tour (available originally in @Buja1986-la rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$. A little tour (available originally in @McDonald1982-ht navigates through all of the coordinate planes,  e.g. in 2D this would correspond to all elements of $A$ are 0 except for one element in each column, $(A_{i1} = A_{j2} = 1, A_{k1} = A_{k2} = 0, ~~i, j, k = 1, ..., p, ~~i\neq j \neq k)$.

A manual tour [@cook_manual_1997] is the most different, because it effectively allows rotating from a projection where variable, $j$, is fully extended in one direction to another where it contributes nothing, i.e. $A[j] \in [0,1]; j=1, ..., p$. The contribution of all other variables is restricted by the initial projection used, and by the contribution of $A[j]$. The initial projection is typically the optimal projection found by the guided tour, and the manual tour is thus used to determine whether which variables are most important to the interesting structure seen.

Figure \ref{fig:notation-target-gen} shows a representation of 1D tour paths of 5D data, drawn on a PCA space. Here PCA provides a convenient 2D representation of the 5D.  A grand tour path is on the left and a guided tour path is on the right. The grand tour path should be more wide-spread because it is attempting to show all possible projections, if left to run long enough. The guided tour path should be short, as the projection pursuit optimization zeroes on the optimally interesting projection. (An interesting aside is that although PCA is used here for creating a summary plot, it can also be considered to be a PP method, where the search finds projections where the data is most varied. There are many different PP indexes that have been proposed over the last few decades. @abdi-williams and @joliffe-cadima-pca provide recent overviews of PCA and @jee2009 provides an overview of PP.)

The recently developed slice tour [@Laa2020-js] can be applied to any of the above tour types. It displays a slice through the orthogonal space, as opposed to a projection, and is explained in Section \@ref(slices).

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/notation-target-gen.pdf}
\caption{A comparison of a guided tour path and grand tour path in a PCA-reduced 2D space. The start point of tour paths is smaller and lighter than the end point. Guided tour optimises the index value iteratively and finishes the search quickly while grand tour wanders in the parameter space to look at possible interesting projections.}
\label{fig:notation-target-gen}
\end{figure}

## Geodesic interpolation

The smooth progression in the tour path is due to geodesic interpolation between the target planes. This takes into consideration two important aspects: (1) it maintains the orthonormality of the projection bases, (2) contains all the rotation to be between planes, not a particular basis in any plane. The first is clearly important because it ensures that we are looking at low-dimensional projections of the data always. The latter is harder to explain, but really important from a visual perspective. It stops any within-plane spin, and could be considered stabilizing the view. A geodesic interpolation between two planes is obtained by (1) using singular value decomposition of $A_1^T A_2$ to compute the $d$ smallest angles between the two planes, (2) dividing these angles into small intervals, (3) generating intermediate projection planes by rotating the initial basis $A_1$ by these angle increments. A more detailed explanation of this can be found in @Buja2005-cx.

Figure \ref{fig:interpolation-2d} shows tour paths of 2D projections of 6D data. The space of all tour paths is a high-dimensional torus, as represented by the gray points. The tour paths are shown in green and orange, and each dot indicates a projection in the sequence.  The plots in this figure could be considered to be a tour looking at itself, because each plot is a selected view from a tour of the torus with the paths overlaid.

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus001.png}
\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus017.png}
\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus064.png}}
\centerline{
\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus068.png}
\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus075.png}
\includegraphics[width=0.3\textwidth]{figures/notation-torus-frames/torus100.png}
}
\caption{A selection of projections from a tour showing the space of possible 2D projections from 6D. Grey points indicate possible 2D projections. The traces of two short paths obtained via geodesic interpolation are shown in color, with starting and ending projections being slightly larger. The space of all 2D projections is a torus, as noted in \cite{Buja1986-sw}. }
\label{fig:interpolation-2d}
\end{figure*}

## Displaying data projections

Various displays can be used to show projections in 1D, 2D and higher dimensions. A 1D projection displays the data in a histogram, while 2D projections are displayed as scatterplots,  analogous to a shadow puppet being projected onto a wall. Cook implemented 3D tours in VRGobi [@vrgobi;@Nelson1999XGobiVT].  Higher dimensional projections can be shown with multivariate displays like SPLOMs or parallel coordinates. Figure \ref{fig:display} shows a variety of displays for projections and these displays are useful to show the non-normal distributions in the projected data, clustering structure, and multivariate relationship between variables. It is possible to define tours for more complex data also. For example, @Wegman98 proposed the image grand tour for multivariate spatial raster data, which displays a linear combination of variables at each spatial pixel. (This is also described in the review paper by @Martinez2013.)

<!-- Figure 5: "Various displays of the projections..." Maybe increasing the axis text size will be good to see the axis labels -->
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/notation-display-1d.pdf}
\includegraphics[width=0.3\textwidth]{figures/notation-display-2d.pdf}
\includegraphics[width=0.3\textwidth]{figures/notation-display-5d.pdf}
\caption{Various displays of the projections: 1D display in histogram (left), 2D display in scatterplot (middle), and 5D display in parallel coordinate plot (right).}
\label{fig:display}
\end{figure}

## Slices, sections and projections {#slices}

Interactive systems like GGobi [@Swayne2003] provide the option of adding sectioning information via linked brushing. Sectioning means we select points that fall in a section of the full parameter space, for example for the purpose of highlighting them in a tour display. This often reveals complementary information, and combining sections and projections can for example provide insights into the dimensionality of a data structure [@Furnas1994-dz].

Sectioning a high-dimensional space in systematic manner (without interactive selection of data sections) is challenging, since there is a lot of freedom in choosing a section. One approach is to define sections based on projection planes, we refer to such sections as "slices" of the data. This is implemented in the slice tour display [@Laa2020-js], and uses the orthogonal distance of each data point from the current projection plane (typically placed such that it passes through the data mean) to highlight points that are nearby and fade out points further from the plane.

The slice tour can reveal concave or non-linear structures obscured in projections, as well as small structures hidden near the center of a distribution. Figure \ref{fig:tour-slice} shows snapshots of the slice tour of points distributed on the surface of geometric shapes. Along with projections, an index function can be used to select target planes to show interesting slices of the data, defined as a section pursuit guided tour [@Laa2020-ni].

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.3\textwidth]{figures/roman-surface-centered-010.png}
\includegraphics[width=0.3\textwidth]{figures/torus-4-centered-015.png}
\includegraphics[width=0.3\textwidth]{figures/cube-6-centered-001.png}}
\centerline{
\includegraphics[width=0.3\textwidth]{figures/roman-surface-centered-032.png}
\includegraphics[width=0.3\textwidth]{figures/torus-4-centered-033.png}
\includegraphics[width=0.3\textwidth]{figures/cube-6-centered-063.png}
}
\caption{A selection of views of a slice tour display showing geometric shapes: Roman surface, 3D (left), 4D torus (middle), and a 6D cube (right). Adapted from \cite{Laa2020-ni}, Figure 4.}
\label{fig:tour-slice}
\end{figure*}

## Transformations on projections

Transformations of data such as scaling, sphering, or a logarithmic transformation for skewed distributions are often useful prior to touring. However, we may also want to transform the data after projecting onto lower dimensions, to correct for unwanted effects of the projections. One example is the piling effect that is observed when projecting a high-dimensional distribution [@Diaconis1984-mv]: projected points are approximately Gaussian in most views, and increasingly concentrating near the center as dimensionality increases. The sage display [@Laa2020-uv] proposes a solution to this issue, via a radial transformation of the projected data. This nonlinear transformation is defined in each point of the projection plane and is sensitive to the overall scale of the data as well as the original dimensionality $p$. In addition, tuning parameters can be used to obtain a more (or less) aggressive redistribution of the points. This is illustrated in Figure \ref{fig:pollen} showing the application of the sage transformation to the classical pollen dataset [@Coleman1986-so]. (This data has five variables and was simulated as a challenge for the 1986 American Statistical Association Data Expo.) The projection without any rescaling looks similar to what is found when rescaling with default options (left), and we can use either of the two tuning parameters to better resolve the distribution near the center and reveal the hidden structure (middle and right).

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/pollen_sage-088.png}
\includegraphics[width=0.3\textwidth]{figures/pollen_sage_R1-088.png}
\includegraphics[width=0.3\textwidth]{figures/pollen_sage_gam20-088.png}
\caption{Snapshots of the same projection of the classical pollen data during a sage tour, where two parameters control a radial transformation of the projected data. Three tuning options are used: (left) default resulting in a similar view as seen in a regular tour; (middle, right) changing the radius and scale parameters, respectively provides better resolution for seeing the hidden word "EUREKA". Adapted from \cite{Laa2020-uv}, Figure 9.}
\label{fig:pollen}
\end{figure}

# Ways of interacting {#interaction}

## Basic interactions

Due to the dynamic nature of the tour, user interfaces can enhance the interpretability of the resulting visualization. Often one will want to pause on a particularly interesting projection and return to use in a downstream analysis or static plot.  Basic controls can be implemented so a tour display can be paused, refreshed and replayed over and over, as shown using the liminal [@Lee2020-ut] R package in Figure \ref{fig:tour-controls}. These basic interactions were available in many historical interactive graphics software, but are available using  liminal or plotly [@sievert_interactive_2020] using projection sequence provided by the tourr [@Wickham2011-uz] R package.

\begin{figure*}
   \centering
         \includegraphics[width=0.9\textwidth]{figures/liminal-stop-pause-refresh.png}
\caption{The clustering example from Figure \ref{fig:tsne-tour} using the liminal interface. Here a grand tour is displayed on the right hand side, with buttons allowing users to play, pause and refresh the tour animation. Adapted from \cite{Lee2020-kg}, Figure 2.}
\label{fig:tour-controls}
\end{figure*}

## Manual tours {#tours}

Manual tours [@cook_manual_1997; @spyrison_spinifex_2020] offer a means to interactively control the contribution of a single variable on the projection plane. This is particularly useful for exploring a projection once a feature of interest has been identified. Manual tours can then be employed to test the structure of the feature, with respect to a selected variable. For instance, Figure \ref{fig:manual-tour}, starts from the orthonormal linear discriminant and explores the sensitivity of class separation as the contribution of a single variable is altered.

\begin{figure*}[!ht]
   \centering
         \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl1_r.pdf}
         \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl3_r.pdf}
         \includegraphics[width=0.32\textwidth]{figures/penguins_manual_bl4_r.pdf}
\caption{Three projections from a manual tour where variable bill length (bl) (shown in orange) from the palmer penguins data is being rotated out of the projection. When this variable is removed, the two light green clusters merge, which informs us that bill length is an important variable for distinguishing between these two species of penguins.}
\label{fig:manual-tour}
\end{figure*}

## Spin-and-brush {#sec:spin}

\begin{figure}[htp]
\centerline{{\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp1.pdf}}
 {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp2.pdf}}
 {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp5.pdf}}}
\smallskip
\centerline{{\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp7.pdf}}
 {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp8.pdf}}
 {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp9.pdf}}}
\centerline{{\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp10.pdf}}
  {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp11.pdf}}
  {\includegraphics[width=0.32\textwidth]{figures/ggobi_book/prim7-pp13.pdf}}}
\caption{Stepwise identification of clusters in the physics data using the XGobi software. Each time a cluster is clearly separated we stop the tour and brush the points, followed by another spin of the data, until all visible clusters are colored. The shape of this 7D data is very interesting: a 2D triangle with two linear clusters extending from each vertex. Drawing lines in a projection helps to outline the triangle shape.  Adapted from \cite{Cook2007-be}, Figure 5.3.}
\label{fig:prim7-tour}
\end{figure}

Brushing can be used to aid statistical and geometrical interpretations of the data; brushing can be thought of as conditioning variables on certain regions of the data or it can be used to section lower dimensional views. This aids tasks such as the identification of outlying points or visual cluster analysis when combined with the tour. Here, a "spin-and-brush" approach works well, since different views will reveal different features and a persistent brush helps us to connect this to the previously observed information. As an example we briefly summarize an analysis from @Cook2007-be that explores clustering of the physics data previously used to in the context of projection pursuit [@Friedman1974-ck; @Cook1995-ae]. This data set is well described by its geometric structure: a two-dimensional triangle with two one-dimensional strands linearly extending in different directions from each vertex. These strands can sequentially be identified as clusters using the spin-and-brush approach, as shown in Figure \ref{fig:prim7-tour}. We stop the tour each time a cluster is clearly separate from the main distribution in the current projection and brush the points in a new color. In the end the full structure becomes apparent and can be visualized by replaying the tour with all clusters highlighted in different colors.

# Software {#software}

There has been a long history of software including tour methods, or something similar. These have laid the foundation for the current tools. The American Statistical Association Statistical Graphics Video Library web site [@ASASGVL] hosts videos documenting the history. It is fabulous watching the videos titled "Multidimensional Scaling" (Kruskal, 1962), "Real-Time Rotation" (Chang, 1970), and "Prim-9" (Tukey, 1973) that some preliminary methods leading to the development of the grand tour. The video titled "Use of the Grand Tour in Remote Sensing" (McDonald and Willis, 1987) is the first to show a tour, and the video titled "Dataviewer: A Program for Looking at Data in Several Dimensions" (Buja and Tukey, 1987) demonstrates the tour as part of a larger data analysis system. The videos titled "XGobi: Dynamic Graphics for Data Analysis" (Swayne et al, 1991) and "Grand Tour and Projection Pursuit" (Cook et al, 1993) show the tour tools in the XGobi software system. Some important works on tours, for example XplorN [@Wegman92] and XLispStat [@Tierney1991], density grand tours [@salch-scott] and computationally efficient grand tours [@Moustafa2010], are not documented in the video collection.

The most accessible current software is the R [@R_Core_Team2020-lr] package called tourr [@Wickham2011-uz]. This software evolved from the Dataviewer [@Buja1986-la], XGobi [@Swayne1998], GGobi [@Swayne2003], Orca [@Sutherland2000], rggobi [@rggobi] and cranvas [@Xie2014] ancestry. The tourr package has a wide range of display types, for different projection dimensions, and a selection of target generating methods including grand, guided, little, local, section, sage and frozen.

The R package spinifex [@spyrison_spinifex_2020] has three primary features. It produces manual tours (predefined path or interactive manipulation, identifies orthonormal global feature bases with the use of the Rdimtools [@you_rdimtools_2020] package. It renders (manual or other) tours as animations exportable to static .gif (gganimate @pedersen_gganimate_2020 package) or interactive .html widgets (plotly @sievert_interactive_2020 package). Lastly it offers an interactive shiny application providing a graphical user interface to quickly explore different variables contribution to structure in the projection.

The liminal R package uses the tour to explore the quality of non-linear dimension reduction algorithms [@Lee2020-kg]. The interface consists of two side-by-side views consisting of a scatter plot displaying a reduced form of the data, and an interactive tour. Controls such as play/stop/restart are implemented allowing a user to pause on interesting projections and return them to their R session for further analysis. Linked brushing is implemented on both views - if  users brush on the scatter plot view they can see if and how an algorithm like t-SNE has distorted distances in higher dimensional space, while if they brush on the tour view, the tour is paused and structures such as cluster separation can be explored.

A Python implementation of the tour written by @Nguyen2020 is also available, and would benefit from further extensions by other developers.

# Applications {#applications}

## Model visualization

\begin{figure*}
  \centering
    \includegraphics[width=0.32\textwidth]{figures/svm-radial-orange.png}
    \includegraphics[width=0.32\textwidth]{figures/svm-radial-green.png}
    \includegraphics[width=0.32\textwidth]{figures/svm-radial-purple.png}
  \caption{A grand tour run on the prediction boundaries of a radial support vector machine. (Left) The prediction region for points classified as orange produces a slice of a cube. (Middle) The prediction region for points classified as green produces another slice of the cube. (Right) The complete prediction region shows the points classified as purple are embedded within the green slice. Adapted from \cite{Wickham2015-do}, Figure 6.}
  \label{fig:svm-tour}
\end{figure*}

@Wickham2015-do espouses the tour as a key component of their approach for visualizing models in data space. They propose using the tour to explore the high dimensional surfaces produced by model fits instead of simple summary statistics.  In Figure \ref{fig:svm-tour} samples have been taken along the prediction region (i.e. the value the model predicts over a grid of observations) of a 3-$d$ support vector machine used to classify three different classes (green, orange purple), by touring over the prediction region we can see how the classifier splits the decision boundaries.

\begin{figure*}
   \centering
         \includegraphics[width=0.9\textwidth]{figures/scheidigger_nn_tour.png}
\caption{A grand tour run on the output of a deep neural network model for classification. The axes correspond to the probabilities that the model of an image belonging to a certain class. Adapted from \cite{Li2020-kg}.}
\label{fig:tour-nn}
\end{figure*}

The grand tour along with direct interaction techniques has been proposed as a technique for understanding the training of deep neural network models [@Li2020-kg]. They proposed a novel method for aligning the output of different layers by interpolating the output of different layers of a deep neural network via the the grand tour. They also apply the grand tour to explore the changes in single layers of the network for each training epoch. For example, by the running tour over the softmax layer outputs of a classification model, an analyst may understand where in the training confusion between classes occur and gain a view of model performance (Figure \ref{fig:tour-nn}). Similarly, the difference between training and test sets can be assessed through side-by-side linked tours of layer outputs. @li_toward_2020 extends this idea by first reducing the dimensionality of output layers with UMAP [@mcinnes2020umap] and then using the grand tour with manual controls over 15 dimensional embeddings to understand layer topology.

## Physics

Physics models often consider $\mathcal{O}(10)$ free parameters and are compared to a much larger number of experimental observables ($\mathcal{O}(100)$). The comparison typically relies on numeric computation of the model predictions for all observables, potentially obscuring the nature of their dependence on the model parameters. Here, multivariate visualization methods can provide new insights. This was demonstrated in @Cook2018-jm. For example, we explored grouping of experiments based on how they constrain the parameter space and we also identified an interesting multivariate outlier that pointed to potential issues with the data point. For illustration we show static views from the tour showing the orthogonal structure of the three groups (indicated by color), and a display which highlights the "outlyingness" of the data point marked with an asterisk symbol in Figure \ref{fig:physics}.

\begin{figure*}
  \centering
           \includegraphics[width=0.45\textwidth]{figures/allcenter-301.png}
         \includegraphics[width=0.45\textwidth]{figures/jetCluster-027.png}
\caption{Selected frames from using a grand tour of the physics data. Left: the three different types of measurements (shown in different colors) are aligned along different directions in parameter space. Right: an outlying point is marked with an asterisk symbol in the ATLAS7new facet. Adapted from \cite{Cook2018-jm}, Figures 6 and 7.}
\label{fig:physics}
\end{figure*}

## Bioinformatics

In single cell RNA sequencing, scientists are interested in identifying novel cell types and understanding the relationships between cells or their developmental trajectory. To achieve this they perform cluster analysis on a counts matrix or principal components and embed the results via t-SNE and label points according the cluster label. One of the main advantages of t-SNE is the avoidance of over-plotting so clusters can be clearly identified on a scatter plot, however, this can come at the cost of interpretability as global distances are distorted.  In @Laa2020-uv, we used radial transformations of the tour projections as an alternative to t-SNE that better preserves global structure while still retaining cluster topology.

\begin{figure*}
  \centering
           \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-001.png}
         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-016.png}
         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-038.png}
         \includegraphics[width=0.24\textwidth]{figures/mouse_sage_2c_gam3-074.png}

\caption{Selected frames from using a grand tour of mouse retina single cell RNA-seq data with the radial transformation. The sage display was used to identify and verify cluster separation between the three highlighted clusters estimated from a clustering algorithm. Adapted from \cite{Laa2020-uv}, Figure 6.}
\label{fig:mouse-cluster}
\end{figure*}

## Geometry of data

High-dimensional data is on many analysts' minds in recent years, and a way to build understanding of high-dimensional spaces is to examine common high-dimensional shapes using a tour. Many different shapes can be simulated using the R package geozoo [@RJ-2016-044]. Figure \ref{fig:geozoo} shows points on the surface of two different types of high-dimensional torii.

\begin{figure*}
  \centering
  \includegraphics[width=0.24\textwidth]{figures/torus-rings-4-with-torus-1.pdf}\includegraphics[width=0.24\textwidth]{figures/torus-rings-4-with-torus-2.pdf}
  \includegraphics[width=0.24\textwidth]{figures/torus-flat-6-1.pdf}\includegraphics[width=0.24\textwidth]{figures/torus-flat-6-2.pdf}
\caption{Several views of torii: (two at left) 4D ring torus, and (two at right) 6D flat torus. In each case, a subset has been highlighted which illustrate the low-dimensional donut shapes. Adapted from \cite{RJ-2016-044}, Figures 10 and 13.}
\label{fig:geozoo}
\end{figure*}

# Discussion {#discussion}

This review highlights the history of tour methods, and modern extensions for the visualization of high dimensional numerical data. The rich ecosystem of tours is being actively developed and applied to a broad range of problems including machine learning. When tours are be combined with some interactivity it can assist with analyses like clustering (the spinifex and liminal packages).  Modified displays of projected data can be used to expose local structures (the sage algorithm, slicing and section pursuit). New algorithm data collection inside the tour code is being used to better diagnose optimization procedures for projection pursuit [@zhang2021visual].

There are many possible future directions for tour research.  One big challenge for software engineering is to seamlessly embed the tour with interactivity. Currently, there is no implementation that analysts can use to perform tasks that were easy in software available in the 80s and 90s. More contributions in sectioning as an alternative to projection, could be broadly useful. An example is conditional model visualization, as implemented in the condviz package [@JSSv081i05], which assists with visual exploration of multivariate fitted models. It could be interesting to define tour paths for different types of nonlinear subspaces, or non-Euclidean space.

# Acknowledgements {-}

The authors gratefully acknowledge the support of the Australian Research Council. The paper was written in `rmarkdown` [@Xie2018-bg] using `knitr` [@Xie2017-uh]. The source material for this paper is available at \url{https://github.com/dicook/wiley-isghdd}.

# References {-}