### 2021-09-01 response_to_review.Rmd
Cleaned up comments on bibliography in response

---
title: "Response to review"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Thank you for the detailed and helpful reviews. These are the ways that we have changed the article in response to the comments.

# Co-editor

- I am fond of this reference, but perhaps Debby's contains it:

Salch, J.D. and D.W. Scott, (1997),
``Data Exploration with the Density Grand Tour,''
 Statistical Graphics and Computing Newsletter,
ASA, Alexandria, VA, 8: 7-11, April, 1997. https://community.amstat.org/jointscsg-section/newsletter

    - **Added**

# Reviewer: 1

- Overall, there is nothing wrong in the article, but lots of language issues, problems with references, and a few aspects of tours (including software and references) that have not been explored in detail. Please see the attached markup for details.

    - **We are extremely grateful for the detailed read and fixes!**

- Before submitting the revision, I would encourage one of the senior native English speaking authors to carefully do a final language review of the article. I have marked many language problems, but there may be more.

## Marked up comments

- check your citation settings; I would have expected that clicking on a citation brings you to that reference in the reference list, but this does not seem to be the case here

    - **This might be a problem with the Wiley template**

- "means" sounds strange; perhaps "refers to the number of dimensions in ..." ?

    - **Changed**

- first introduce "one-dimensional (1D) through to five-dimensional (5D)"; then you can use 1D, ..., 5D

    - **Changed**

- perhaps "As the dimension increases, ..."

    - **Changed**

- change to  "a 2D cube consists of ..." and "a 3D cube consists of ..."

    - **Changed**

- fig 1: explain how the cube doubles: by connecting the vertexes of the two cubes. can you use 3 colors: two colors for the two original cubes and a third color for the connecting edges of the vertexes; perhaps draw those connectors as dashed lines

    - **Done**

- p2 l49: perhaps "seen that as the dimension increases by one, ..."

    - **Done**

- p3 l14 just "e.g.," or "see for example"

    - **Done**

- p3 l14 mention the underlying software as well (and then cite these references)

    - **Done**

- p3 l16 what are geodesic interpolations? explain in 1 or2 sentences and add a reference

    - **Done**

- p3 l17 "a multivariate ..." or "distributions"

    - **Done**

- p3 l28 this sounds like a limitation of PCPs on your side, but Hurley & Oldford found a possible solution, e.g., https://link.springer.com/article/10.1007/s00180-011-0229-5

    - **added a sentence and the reference**

- p3 l28 you mention Andrews curves later on when you speak about software; mention them here as well and clarify that they are some kind of tour, but not a grand tour; Wegman & Solka (2002) describe Andrews plot as a pseudo-grand tour that is not space filling - see
On Some Mathematics for Visualizing High Dimensional Data
Author(s): Edward J. Wegman and Jeffrey L. Solka
Source: Sankhyā: The Indian Journal of Statistics, Series A (1961-2002) , Jun., 2002, Vol.
64, No. 2, Selected Articles from San Antonio Conference in Honour of C. R. Rao (Jun.,
2002), pp. 429-452
Published by: Indian Statistical Institute
Stable URL: https://www.jstor.org/stable/25051404

    - "Added, and explained"

- p3 l28 Wegman & Solka (2002) also provide mathematical details and further references; while it is not necessary for this article to repeat all of that information, it should at least summarize those mathematical references in one paragraph.

    - "Not done. We have reference the paper but the Buja et al is the better reference for mathematics."

- p3 l32 "analyst" (singular)

    - **Done**

- p3 l43 be specific which reference is for which of the 2 techniques

    - **Done**

- p3 l51 add "have" ?

    - **Done**

- p4 l27 do not just show some graphs; provide at least some meaningful background information: 2 species of crabs where 0 represents xxx and 1 represents yyy; 2 sexes (which is M & F); the data in the left plot seem to be jittered , etc.; what are dimensions in the right plot - cm?; also cite the original article where these data have first been reported and analyzed

    - **Done**

- p4 l41 add a "," here

    - **Done**

- p5 fig 3 what does flat and hierarchical mean here?

    - **removed legend because these labels are not really sensible or useful**

- p5 fig3 be consistent and write as "10D"

    - **Done**

- p5 fig3 perhaps "can be seen" or "shown"

    - **used "visible"**

- p5 fig3 perhaps "a sample" or "a selection" (as you picked specific projections)

    - **Done**

- p5 fig3 distance does not seem to fit here - perhaps "distant"

    - **Done**

- p5 l38 add "a"

    - **Done**

- p5 l40 perhaps "all columns are..."

    - **Done**

- p5 2.2 header these are rarely used; briefly explain what a Huber plot is and how it can be interpreted

    - **Done**

- p5 l43 as for the two previous data examples, briefly explain the data set, source, variables, and other information; what are the dimensions of this data set?

    - **Done**

- p5 l48 "targets" (plural)

    - **Done**

- p5 l49 "projections"

    - **Sentence changed**

- p5 l50 cite 1 or 2 references for projection pursuit, including the WIREs article - see
https://wires.onlinelibrary.wiley.com/doi/abs/10.1002/wics.23

    - **Introduced in section 1 where PCA and LDA is first mentioned**

- p5 l53 define PCA and include references; one should be the WIREs article - see
https://onlinelibrary.wiley.com/doi/10.1002/wics.101

    - **added more details on both PCA and PP here**

- fig 4 what does "best.proj.data" represent - and what does the scale for this variable mean?

    - **changed**

- p7 l35 perhaps "these two tour types." ?

    - **Paragraph is now changed**

- p7 l37 be specific: "in Section xxx"

    - **done**

- p7 l39 same: "in Section yyy"

    - **Paragraph changed and it doesn't fit to forward reference here**

- p7 l42 perhaps "it maintains" or "to maintain"; same for (2) below

    - **done**

- p7 l44 "rotations"

    - **singular is correct here**

- p7 l45 "are" not "a"

    - **done**

- p7 l46 "projections{

    - **done**

- p7 l47 "A more detailed ..."

    - **done**

- fig 6 caption "paths", "The guided...", "the grand..."

    - **done**

- p8 l30 Cook worked with 3D tours in VRGobi; ideally cite one of those original references where the 3D tour is introduced or cite this chapter (in particular Section 10.5.2) from a book chapter by Symanzik as a secondary reference:
http://sfb649.wiwi.hu-berlin.de/fedc_homepage/xplore/ebooks/html/csa/node83.html
This is from 
Handbook of Computational Statistics by	 
J.E. Gentle, W. Härdle, Y. Mori, Editors, ISBN-10: 3540404643

    - **done**

- fig 7 caption remind the reader that small points are identical start points while large points are different end points AND emphasize that each of these 6 views shows exactly the same tour, just under a different projection in 2D space

    - **done**

- p9 l36 change order; perhaps "Transformations of data (such as ..) are often useful prior to touring.

    - **done**

- p9 l45 as before, briefly summarize this data set (dimensions, number of points, and the text that magically occurs with the proper methods are being applied)

    - **done**

- fig 9 caption how many dimensions does this have?; add ","; "6D"

    - **done**

- fig 10 caption what do R and gamma tune? can you verbally explain their purpose?

    - **done**

- section 3 header be consistent with capitalizations of headings, e.g., lower case "i" here but upper case in the subheading below (check and adjust elsewhere)

    - **done**

- p12 l11 add "to"

    - **done**

- p12 l12 which software is this? and which options for tour interactions are commonly available in other software with tour functionality such as XGobi, RGobi, tourr ?

    - **This has been done as much as possible, and the reference to rggobi is added**

- p12 l25 "or it can be used"

    - **done**

- p12 l30 add "set"

    - **done**

- fig 12 caption perhaps "variable "bl" (shown as an orange line) is being..."; as before, describe the data set, source, dimensions, variables (what is bl), etc.?; add ","

    - **done**

- fig 13 what is hidden (and why)?; what is the purpose of "connect dots"? how are they connected (this does not look like the convex hull in the projection plane); which software has been used here?

    - **done**

- p15 l11 add ","

    - **done**

- p15 l11 add "that"

    - **done**

- p15 l21 indicate the names of the software, i.e., XplorN and (X)Lisp-Stat; in particular, XplorN deserves some further description as it was one of the first software implementations that provided a PCP grand tour; here is a suitable technical report (follow the cite references for more):
https://apps.dtic.mil/sti/pdfs/ADA313999.pdf

    - **done**

- p15 l22 this paragraph doesn't seem to belong here as it doesn't seem to relate to software; I suggested earlier on that Andrews curves should be introduced earlier in this review; this paragraph can be used there

    - **Moved to the intro with other discussion of lietrature**

- p15 l26 explain in one sentence what the image grand tour is and cite one original article for it (likely something from Wegman et al); by the way, here, you cite the WIREs review article, but an original article is missing; for several of my previous comments, you cited the original article, but not the WIREs review article; ideally, both (original and review article where it exists) should be cited in a review like this

    - **done**

- p15 l28 also cite rggobi as it seemed to support some tour as well - see https://vita.had.co.nz/papers/rggobi-rnews.pdf

    - **done**

- p15 l33 write as "three"

    - **done**

- p15 l36 sentence structure is too complex; break at least into 2 sentences and avoid the double parentheses for (software (author))

    - **done**

- p15 l44 remove "a"

    - **done**

- p15 l46 perhaps "structures such as .. separations"

    - **done**

- p15 l48 perhaps "from further extensions" ?

    - **done**

- fig 14 make sure to indicate which software has been used for each tour figure; as before, provide more details about the underlying data; Be consistent how to introduce multiple plots in the same figures; you use (left), (Left), Left: etc.

   - **done**

- p16 l31 3D (for consistency)

   - **done**

- p16 l32 add ","

   - **done**

- p16 l32 period here; then start a new sentence: "By touring..."

   - **done**

- p16 l35 you mix present tense and past tense when you summarize the content of references; here, you use past tense, elsewhere, you use present tense; at least, be consistent; personally, I prefer past tense as authors did something in the past (and not in the present)

   - **fixed**

- p16 l37 perhaps "running the" ?

   - **done**

- p16 l38 add "the"

   - **done**

- p16 l47 do you need the O notation here? what about dozens and hundreds instead?

   - **done**

- p16 l50 "we" or "they" (several times)? [and note that this is past tense again]

   - **done**

- fig 15 as always, much more details about the underlying data needed

   - **done**

- p17 l36 add ","

   - **done**

- p17 l41 again "we"; perhaps "the authors"

   - **done**

- p17 l41 mention Figure 17 here in the text

   - **done**

- p17 add "5.5 Historical applications" and briefly review other examples from the literature here; e.g., Koschat and Swayne seem to have used the tour for telecommunication data (see below)

    - **Phew! Tall order. Added a few examples to illustrate the range of applicability**

- fig 16 as always. more details about the data needed

    - **done**

- fig 16 what is ATLAS7new facet ? explain in main text

    - **this was a typo - fixed**

- fig 18 perhaps "subset... illustrates ... shape"

   - **done**

- p19 l10 remove "be"

   - **done**

- p19 l12 add "R" (as you mentioned Python and other software before)

   - **done**

- p19 l12 "it" ? - perhaps "supporting software" ?

   - **done**

- p19 l13 add ","

   - **done**

- p19 add a paragraph about success of the various tour methods; Cook et al (1995) stated this (p.163): 
"but it is important to statisticians for the reason that the inherent structure has never been completely described. The combination of the grand tour and projection pursuit contributes significantly to revealing the nature of the variable relationships in seven dimensions. Recently, Koschat and Swayne (1992, in press) have used the projection pursuit guided tour in XGobi to explore telecommunications data, and indeed found previously undetected structure. "

   - **done**

- p19 l18 add "to perform"

   - **done**

- p19 l19 add ","

   - **done**

- p19 l19 be specific what is missing in current software, i.e., continue with "(such as ...)."

   - **done**

- p19 l23 perhaps "spaces" (there could be more than just one?)

   - **done**

- References
    - video library url needed - done
    - buja et al in wegman & solka add city - done
    - buja et al (1986) this is a conference proceedings series; capitalize similar to a journal, add page numbers - this probably needs to be fixed via bibstyle (no page numbers showing for proceeding entries)
    - chambers et al list the editors - it seems there was no relevant editor to list here
    - coleman add when accessed - date added based on animation package (data was accessed through the package)
    - cook et al (1995) adjust style (compare with previous reference) - done
    - cook et al (2018) remove the ArXiv information for an article that has been published - done
    - cook, swayne (2007) reference is wrong add add city of publication (New York, Berlin, Heidelberg?) - done
    - diaconis as above - either full journal names or abbreviations (adjust other references in a similar way, according to WIREs specifications) - done
    - fisher remove the first URL and just keep the one with the DOI - done
    - friedman is there a DOI? - added
    - horst capitalize both "A" as these seem to be geographic names; just "R package" once - fixed capitalization, but repetition needs to be fixed via bibstyle
    - hotelling should be "417-441" - fixed (plus full journal name instead of abbreviation)
    - huber is there a URL where this can be obtained? - unfortunately not
    - laa et al (2020) section pursuit incomplete - fixed
    - laa et al (2021) burning sage incomplete - fixed
    - laa et al (2020) slice tour update reference (it is published now) and remove the arXiv information - fixed
    - lee at al DOI ? - added
    - li incomplete; use lower case "c" and "t" here - fixed
    - li where does the 5(3) come from? - check the web page again - fixed
    - martinez remove the first URL - fixed
    - mcinnes capitalize "UMAP" - fixed
    - moustafa incomplete; Remove the first URL - fixed
    - moustafa and hadi remove the first URL - fixed
    - nguyen capitalize P (in Python), R (in the next reference), etc. - fixed
    - pedersen just once; check gganimate capitalisation - fixed capitalisation, but repeating version number is bibstyle issue
    - sievert city?; should be "Chapman and Hall/CRC" (several times) - fixed
    - spyrison change to lower case (except R); 243-257 (again!) - fixed
    - sutherland et al DOI - added
    - swayne DOI - information is in the bib file but not showing (maybe bibstyle issue?)
    - swayne et al DOI; XGobi (as a software name); issue ? -fixed
    - tukey editors? - added
    - tukey pages, city, publisher ?; city? - added
    - van der maaten they list this as 9(86) on their web page - fixed
    - wegman list editors only once; page numbers; is there a volume ? - these are Interface (conference proceedings) - problem with showing these information for proceedings entry, again this appears to be a bibstyle issue
    - wilkinson capital "S" - fixed
    - xie lower case - fixed
    - you "R" - fixed
    - zhang extend reference as for other R packages; incomplete - fixed

    - **Fixed. Some of the problems are related to the template,. The bibtex files are complete so we hope that the typesetter can suitably make the complete citations.**

- I am curious: Why do you not cite https://doi.org/10.1051/epjconf/202024506018

    - **This was a conference presentation and summarised work in other papers. It looks like a paper, which is the physics approach, but it cannot be considered a paper in its own right.**

- as you cite R packages, reference R as well

    - **done**

# Reviewer 2

## Summary

- This work summarizes the state-of-the-art tour methods for visualizing high-dimensional data. The tour works by linearly projecting the set of data points down to low dimensions through a sequence of smoothly changing orthonormal projections. The work summarizes the development, components for the display, interaction modes, software and latest applications of the tour method. Overall, the work is well-written and well-covered.  I would recommend a minor revision to authors sometime to revise the content and text (see recommended revisions).

    - **Thanks for the positive comments**

## Strengths

- The literature is well covered and described, and up to date. There are abundant number of examples with figures which covers a large variety of use cases of the tour method (e.g. kinds of display, structure of measurements in physics, geometric shapes, a tour looking at itself, ...)

    - **Thanks for the positive comments**

## Recommended Revisions

- I highly appreciate the specification of notations in Section 2.1, but few sections use the notation.  Instead, different tour methods are only described in text. 
I anticipate that the readers would benefit from a highlight of key differences between different tour variants and interactions using the unified notations. For example, (as far as I understand) guided tour finds a sequence of interesting projections $A_i$ by applying projection pursuit optimization on a predefined criteria $c(\cdot)$ of ``interestingness'' of a projection, $argmax_A c(XA)$; local tour rocks back and forth from a particular plane $A$ to randomly chosen targets $A_i$ in a small neighborhood $\{A_j |\; d(A_j, A) < \epsilon \}$; manual tour manipulates the contribution of any variable by editing the corresponding row $A[i]$ in the projection frame $A$, $A[i] \leftarrow A[i] + \Delta[i]$ and re-orthonormalizing the columns.

    - **Done**

- Section 2.2: "A guided tour ... The little tour ... and a local tour ..." --- cite?

    - **Done, cited as much as is possible. There isn't a good citation for little tour or local tour, so we have referred to the original Dataviewer software.**

- Section 2.3: "contains all the rotation to be between planes, not a particular basis in any plane". It maybe worth mentioning the difference between "interpolating frames" (interpolating projection frame A and B directly, which involves particular bases) vs. "interpolating planes" (between span(A) and span(B) ), and briefly discuss how the motion can be decomposed to motion of planes and within plane spin.

    - **done**

- Section 2.3: "The space of all tour *paths* is a high-dimensional torus..." --- All paths, all k-planes or all orthogonal k-frames? 
From the context it's not clear to me whether the space of all tour paths is a high-D torus. 
Consider giving some contexts/explanations to this statement?

    - **cited the original buja and asimov paper that states this, and you can see it from the torus shapes in Figure 18, too.**

- Section 3: Li et al. 2020 seems to propose a different way to control the projection, via brushing and dragging data points as opposed to moving individual variables in manual tour (what they call ``the data point mode'' direct manipulation). 
Perhaps it is worth mentioning it as a way of manual control in Section 3.

    - **done**